\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{physics}
\usepackage{tikz}
\usepackage{float}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\title{Arrowhead Matrix and Berry Phase Calculations: \\
       Technical Documentation}
\author{Arrowhead Project Team}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This document provides a comprehensive technical overview of the Arrowhead matrix generation, Berry phase calculations, and visualization tools implemented in the project. We focus particularly on the \texttt{run\_optimal\_visualization.py} script, which generates comprehensive visualizations of Berry phases, eigenstate behavior, and parity flips using optimal parameters. We also discuss the physical significance of half-integer winding numbers and topological phase transitions observed in the system.
\end{abstract}

\tableofcontents

\section{Introduction}

The Arrowhead project implements and analyzes a quantum system described by Arrowhead matrices, with a focus on Berry phase calculations and topological properties \cite{Xiao2010, Vanderbilt2018}. The project includes tools for:

\begin{itemize}
    \item Generating Arrowhead matrices with configurable parameters
    \item Calculating Berry phases with improved eigenstate tracking
    \item Visualizing Berry phases, eigenstate behavior, and parity flips
    \item Analyzing topological phase transitions
\end{itemize}

This document provides a detailed explanation of the key components, with a focus on the \texttt{run\_optimal\_visualization.py} script and its role in analyzing and visualizing the system's behavior.

\section{Arrowhead Matrix Generation}

\subsection{Mathematical Background}

An Arrowhead matrix is a special type of matrix with non-zero elements only in the first row, first column, and main diagonal. In our implementation, we generate a $4 \times 4$ Arrowhead matrix with specific parameter dependencies.

The general form of a $4 \times 4$ Arrowhead matrix is:

\begin{equation}
A = \begin{pmatrix}
a & b_1 & b_2 & b_3 \\
b_1 & c_1 & 0 & 0 \\
b_2 & 0 & c_2 & 0 \\
b_3 & 0 & 0 & c_3
\end{pmatrix}
\end{equation}

In our implementation, these elements depend on various parameters including $\theta$ (angle), $x\_shift$, $y\_shift$, $d\_param$, $\omega$, $a\_vx$, and $a\_va$.

\subsection{Implementation}

The Arrowhead matrix generation is implemented in the \texttt{generate\_4x4\_arrowhead.py} script. The key steps are:

\begin{enumerate}
    \item Define the parameter space, including $\theta$ values ranging from 0 to 360 degrees
    \item Generate orthogonal R-vectors based on $\theta$
    \item Calculate the VX and VA potentials for each component of the R-vector
    \item Construct the Arrowhead matrix using these potentials
    \item Save the matrix to a file for later analysis
\end{enumerate}

\subsubsection{Potential Functions}

The potentials are implemented as parabolic functions:

\begin{equation}
V_X(\mathbf{R}) = a \cdot ||\mathbf{R}||^2 + c
\end{equation}

where $a = 0.05$ is the curvature parameter, $||\mathbf{R}||$ is the magnitude of the position vector, and $c = 0$ is a constant offset.

\begin{equation}
V_A(\mathbf{R}) = a \cdot ||\mathbf{R} - \mathbf{R}_{shift}||^2 + c
\end{equation}

where $a = 0.2$ is the curvature parameter, $\mathbf{R}_{shift} = (x\_shift, y\_shift, 0)$ with $x\_shift = 22.5$ and $y\_shift = 567.72$ representing the spatial shift of the potential, and $c = 0$ is a constant offset.

\subsubsection{Matrix Construction}

The R-vector is decomposed into three components:
\begin{align}
\mathbf{R}_0 &= (R_x, 0, 0) \quad \text{(x component)} \\
\mathbf{R}_1 &= (0, R_y, 0) \quad \text{(y component)} \\
\mathbf{R}_2 &= (0, 0, R_z) \quad \text{(z component)}
\end{align}

The potentials are calculated for each component:
\begin{align}
v_{x0} &= V_X(\mathbf{R}_0) \quad v_{x1} = V_X(\mathbf{R}_1) \quad v_{x2} = V_X(\mathbf{R}_2) \\
v_{a0} &= V_A(\mathbf{R}_0) \quad v_{a1} = V_A(\mathbf{R}_1) \quad v_{a2} = V_A(\mathbf{R}_2)
\end{align}

The diagonal elements of the matrix are then constructed as:
\begin{align}
D_{00} &= v_{x0} + v_{x1} + v_{x2} + \hbar\omega \\
D_{11} &= v_{a0} + v_{x1} + v_{x2} \\
D_{22} &= v_{x0} + v_{a1} + v_{x2} \\
D_{33} &= v_{x0} + v_{x1} + v_{a2}
\end{align}

The off-diagonal elements (coupling constants) are set to a fixed value $c_1 = c_2 = c_3 = 0.1$.

\section{Berry Phase Calculation}

\subsection{Theoretical Background}

The Berry phase is a geometric phase acquired by a quantum state as it evolves adiabatically around a closed loop in parameter space \cite{Bernevig2013, Hasan2010}. For a quantum state $|\psi(\theta)\rangle$ evolving with parameter $\theta$, the Berry phase is given by:

\begin{equation}
\gamma = i \oint \langle \psi(\theta) | \nabla_\theta | \psi(\theta) \rangle d\theta
\end{equation}

In discrete form, for a path divided into $N$ points, the Berry phase can be calculated as:

\begin{equation}
\gamma = -\text{Im} \log \prod_{j=1}^{N} \langle \psi(\theta_j) | \psi(\theta_{j+1}) \rangle
\end{equation}

Alternatively, this can be expressed as a sum:

\begin{equation}
\gamma = -\text{Im} \sum_{j=1}^{N} \log \langle \psi(\theta_j) | \psi(\theta_{j+1}) \rangle
\end{equation}

Or in terms of the gradient of the wavefunction:

\begin{equation}
\gamma = \text{Im} \sum_{j=1}^{N} \langle \psi(\theta_j) | \nabla_\theta | \psi(\theta_j) \rangle \Delta\theta_j
\end{equation}

These formulations are equivalent for a discretized path in parameter space, with $\theta_{N+1} = \theta_1$ to ensure a closed loop.

\subsection{The Role of the Logarithm in Berry Phase Calculations}

The logarithm in the Berry phase formula serves a crucial mathematical purpose \cite{Xiao2010, Vanderbilt2018, Resta2000}:

\begin{enumerate}
    \item \textbf{Complex Inner Products}: The inner product $\langle \psi(\theta_j) | \psi(\theta_{j+1}) \rangle$ yields a complex number with both magnitude and phase information.
    
    \item \textbf{Phase Accumulation}: As we traverse a closed loop in parameter space, we need to accumulate the phase changes between consecutive states.
    
    \item \textbf{Multiplicative Nature of Phases}: In quantum mechanics, phases multiply rather than add directly. When consecutive transformations have phases $e^{i\phi_1}$ and $e^{i\phi_2}$, the total phase is $e^{i(\phi_1+\phi_2)}$ \cite{Griffiths2017}.
    
    \item \textbf{Converting Multiplication to Addition}: The logarithm converts this multiplication into addition through the property $\log(z_1 \cdot z_2 \cdot ... \cdot z_n) = \log(z_1) + \log(z_2) + ... + \log(z_n)$.
    
    \item \textbf{Extracting the Phase}: For a complex number $z = re^{i\phi}$, we have $\log(z) = \log(r) + i\phi$, so $\text{Im}[\log(z)] = \phi$ \cite{Bohm2003}.
\end{enumerate}

This explains why the two formulations are equivalent:

\begin{align}
\gamma &= -\text{Im} \log \prod_{j=1}^{N} \langle \psi(\theta_j) | \psi(\theta_{j+1}) \rangle \\
&= -\text{Im} \sum_{j=1}^{N} \log \langle \psi(\theta_j) | \psi(\theta_{j+1}) \rangle
\end{align}

The logarithm converts the product of inner products (representing overlaps between consecutive states) into a sum of phases, and the imaginary part extracts just the phase information needed for the Berry phase \cite{Resta2000, Bohm2003}.

The winding number $W$ is related to the Berry phase by \cite{Park2011, Liang2015}:

\begin{equation}
W = \frac{\gamma}{2\pi}
\end{equation}

\subsection{Improved Berry Phase Calculation}

Our implementation uses an improved Berry phase calculation algorithm with eigenstate tracking, implemented in \texttt{run\_improved\_berry\_phase.py}. Key features include:

\begin{enumerate}
    \item Loading Arrowhead matrices for different $\theta$ values
    \item Calculating eigenstates and eigenvalues
    \item Tracking eigenstates across different $\theta$ values to ensure continuity
    \item Calculating Berry phases and winding numbers
    \item Detecting and analyzing parity flips
\end{enumerate}

The eigenstate tracking is crucial for correctly calculating Berry phases, especially when eigenstates cross or become degenerate.

\section{Optimal Visualization Script}

The \texttt{run\_optimal\_visualization.py} script is a comprehensive tool for visualizing and analyzing the results of Berry phase calculations using optimal parameters.

\subsection{Script Overview}

The script performs the following key functions:

\begin{enumerate}
    \item Runs a Berry phase calculation with optimal parameters
    \item Extracts and processes the results
    \item Generates visualizations of Berry phases, eigenstate behavior, and parity flips
    \item Creates a comprehensive summary file
\end{enumerate}

\subsection{Optimal Parameters}

The script uses the following optimal parameters, which were determined through systematic parameter exploration:

\begin{verbatim}
x_shift: 22.5
y_shift: 547.7222222222222
d_param: 0.005
omega: 0.025
a_vx: 0.018
a_va: 0.42
\end{verbatim}

These parameters result in zero parity flips for eigenstate 3, while maintaining the expected topological properties of the system.

\subsection{Visualization Components}

The script generates several types of visualizations:

\begin{itemize}
    \item \textbf{Berry Phase Plots}: Bar charts showing Berry phases for each eigenstate
    \item \textbf{Parity Flip Plots}: Visualizations of parity flips for each eigenstate
    \item \textbf{Eigenstate vs. $\theta$ Plots}: Plots showing how eigenstate values change with $\theta$
    \item \textbf{Eigenstate Degeneracy Plots}: Analysis of eigenstate degeneracy and crossings
    \item \textbf{Potential Plots}: Visualizations of $V_X$ and $V_A$ potentials
    \item \textbf{Comprehensive Infographic}: A combined visualization showing all key aspects
\end{itemize}

\subsection{Summary File Generation}

The script generates a comprehensive summary file (\texttt{summary.txt}) that includes:

\begin{itemize}
    \item Parameter values
    \item Berry phase table with raw phases, winding numbers, normalized and quantized phases
    \item Parity flip summary
    \item Analysis of eigenstate 2's half-integer winding number
    \item Eigenvalue normalization information
    \item Eigenstate degeneracy analysis
\end{itemize}

\subsection{Half-Integer Winding Numbers}

A key feature of the script is its correct handling of half-integer winding numbers, particularly for eigenstate 2. When the normalized Berry phase is close to $\pm\pi$, the script correctly identifies and displays a half-integer winding number ($-0.5$).

This is physically significant because \cite{Tse2010, Hsieh2009, Zeng2024}:

\begin{itemize}
    \item Half-integer winding numbers indicate non-trivial topology
    \item They correspond to Berry phases of $\pm\pi$
    \item They are associated with high numbers of parity flips (129 for eigenstate 2)
\end{itemize}

\section{Topological Phase Transitions}

\subsection{Phase Transition Analysis}

The project includes tools for analyzing topological phase transitions as system parameters are varied \cite{Moore2010, Qi2010, Wen2017, Chiu2016}:

\begin{itemize}
    \item \texttt{run\_parameter\_sweep.py}: Runs simulations with different parameter values
    \item \texttt{plot\_phase\_transitions.py}: Creates visualizations of phase transitions
\end{itemize}

\subsection{Visualization of Phase Transitions}

The visualization of phase transitions helps identify critical points where the system's topological properties change \cite{Shen2017, Asboth2016}.

The phase transition visualizations show:

\begin{itemize}
    \item How Berry phases change with parameter values
    \item How winding numbers change with parameter values
    \item Highlighted transition regions where winding numbers change
    \item Detailed analysis of eigenstate 2's behavior across transitions
\end{itemize}

\subsection{Physical Significance}

Topological phase transitions are characterized by changes in winding numbers and correspond to fundamental changes in the system's topological properties. These transitions are analogous to quantum phase transitions and provide insights into the system's behavior under parameter variations.

\section{Code Implementation Details}

\subsection{run\_optimal\_visualization.py}

The \texttt{run\_optimal\_visualization.py} script is structured as follows:

\begin{lstlisting}[language=Python, caption=Key components of run\_optimal\_visualization.py]
def run_berry_phase_calculation(x_shift, y_shift, d_param, omega, a_vx, a_va, theta_step=1):
    """Run Berry phase calculation with the given parameters."""
    # Implementation details...

def extract_eigenstate_vs_theta_data(results_file):
    """Extract eigenstate vs theta data from results file."""
    # Implementation details...

def create_summary_file(results_file, output_file, eigenstate_data):
    """Create a summary file with Berry phase results and analysis."""
    # Implementation details...
    
    # Berry phase table formatting with proper tabulation
    # Handling of half-integer winding numbers
    # Detailed explanation of eigenstate 2's behavior
    
    # Implementation details...

def main():
    """Main function to run optimal visualization."""
    # Set optimal parameters
    x_shift = 22.5
    y_shift = 547.7222222222222
    d_param = 0.005
    omega = 0.025
    a_vx = 0.018
    a_va = 0.42
    
    # Run Berry phase calculation
    # Generate visualizations
    # Create summary file
    # Implementation details...
\end{lstlisting}

\subsection{Berry Phase Visualization}

The Berry phase visualization components are implemented in \texttt{berry\_phase\_visualization.py}:

\begin{lstlisting}[language=Python, caption=Key visualization functions]
def plot_berry_phases(berry_phases, output_dir):
    """Plot Berry phases for each eigenstate."""
    # Implementation details...

def plot_parity_flips(parity_flips, output_dir):
    """Plot parity flips for each eigenstate."""
    # Implementation details...

def plot_eigenstate_vs_theta(eigenstate_data, output_dir):
    """Plot eigenstate values vs theta for all eigenstates."""
    # Implementation details...

def plot_eigenstate_degeneracy(eigenstate_data, output_dir):
    """Plot the degeneracy between eigenstates."""
    # Implementation details...

def plot_phase_transition(parameter_values, berry_phases, winding_numbers, output_dir, param_name='y_shift'):
    """Plot phase transition as a function of a system parameter."""
    # Implementation details...
\end{lstlisting}

\section{Results and Findings}

\subsection{Berry Phase Analysis}

The Berry phases for the four eigenstates show distinct patterns \cite{Xiao2010, Vanderbilt2018}:

\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
Eigenstate & Raw Phase (rad) & Winding Number & Normalized Phase & Quantized Phase \\
\hline
0 & 0.000000 & 0 & 0.000000 & 0.000000 \\
1 & 0.000000 & 0 & 0.000000 & 0.000000 \\
2 & 0.000000 & -0.5 & -3.141593 & -3.141593 \\
3 & 0.000000 & 0 & 0.000000 & 0.000000 \\
\hline
\end{tabular}
\caption{Berry phases and winding numbers for each eigenstate}
\end{table}

\subsection{Parity Flip Analysis}

Parity flips provide another perspective on the system's behavior:

\begin{table}[H]
\centering
\begin{tabular}{|c|c|}
\hline
Eigenstate & Parity Flips \\
\hline
0 & 0 \\
1 & 13 \\
2 & 129 \\
3 & 0 \\
\hline
\end{tabular}
\caption{Parity flips for each eigenstate}
\end{table}

\subsection{Topological Phase Transitions}

By varying the y\_shift parameter, we observed topological phase transitions characterized by changes in winding numbers \cite{Liang2015, Park2011}. These transitions provide valuable insights into how the system's topological properties depend on its parameters.

\section{Conclusions}

The Arrowhead matrix and Berry phase calculations implemented in this project reveal rich topological properties, including half-integer winding numbers and topological phase transitions. The \texttt{run\_optimal\_visualization.py} script provides comprehensive tools for visualizing and analyzing these properties.

Key achievements include:

\begin{itemize}
    \item Correct identification and display of half-integer winding numbers
    \item Comprehensive visualization of Berry phases, eigenstate behavior, and parity flips
    \item Analysis of topological phase transitions
    \item Optimization of parameters to achieve zero parity flips in eigenstate 3
\end{itemize}

These tools and methodologies provide a solid foundation for further exploration of topological properties in quantum systems.

\section{References}

\begin{thebibliography}{99}

\bibitem{Tse2010} W.-K. Tse and A. H. MacDonald, ``Giant magneto-optical Kerr effect and universal Faraday effect in thin-film topological insulators,'' Physical Review Letters, vol. 105, no. 5, p. 057401, 2010.

\bibitem{Hsieh2009} D. Hsieh et al., ``Observation of unconventional quantum spin textures in topological insulators,'' Science, vol. 323, no. 5916, pp. 919-922, 2009.

\bibitem{Moore2010} J. E. Moore, ``The birth of topological insulators,'' Nature, vol. 464, no. 7286, pp. 194-198, 2010.

\bibitem{Qi2010} X.-L. Qi and S.-C. Zhang, ``The quantum spin Hall effect and topological insulators,'' Physics Today, vol. 63, no. 1, pp. 33-38, 2010.

\bibitem{Park2011} C.-H. Park and N. Marzari, ``Berry phase and pseudospin winding number in bilayer graphene,'' Physical Review B, vol. 84, no. 20, p. 205440, 2011.

\bibitem{Liang2015} S.-D. Liang and G.-Y. Huang, ``Topological invariance and global Berry phase in non-Hermitian systems,'' Physical Review A, vol. 87, no. 1, p. 012118, 2015.

\bibitem{Zeng2024} X.-L. Zeng, W.-X. Lai, Y.-W. Wei, and Y.-Q. Ma, ``Quantum geometric tensor and the topological characterization of the extended Su-Schrieffer-Heeger model,'' Chinese Physics B, 2024.

\bibitem{Vanderbilt2018} D. Vanderbilt, ``Berry Phases in Electronic Structure Theory: Electric Polarization, Orbital Magnetization and Topological Insulators,'' Cambridge University Press, 2018.

\bibitem{Bernevig2013} B. A. Bernevig and T. L. Hughes, ``Topological Insulators and Topological Superconductors,'' Princeton University Press, 2013.

\bibitem{Xiao2010} D. Xiao, M.-C. Chang, and Q. Niu, ``Berry phase effects on electronic properties,'' Reviews of Modern Physics, vol. 82, no. 3, p. 1959, 2010.

\bibitem{Asboth2016} J. K. Asbóth, L. Oroszlány, and A. Pályi, ``A Short Course on Topological Insulators: Band Structure and Edge States in One and Two Dimensions,'' Springer, 2016.

\bibitem{Hasan2010} M. Z. Hasan and C. L. Kane, ``Colloquium: Topological insulators,'' Reviews of Modern Physics, vol. 82, no. 4, p. 3045, 2010.

\bibitem{Wen2017} X.-G. Wen, ``Colloquium: Zoo of quantum-topological phases of matter,'' Reviews of Modern Physics, vol. 89, no. 4, p. 041004, 2017.

\bibitem{Chiu2016} C.-K. Chiu, J. C. Y. Teo, A. P. Schnyder, and S. Ryu, ``Classification of topological quantum matter with symmetries,'' Reviews of Modern Physics, vol. 88, no. 3, p. 035005, 2016.

\bibitem{Resta2000} R. Resta, ``Manifestations of Berry's phase in molecules and condensed matter,'' Journal of Physics: Condensed Matter, vol. 12, no. 9, p. R107, 2000.

\bibitem{Griffiths2017} D. J. Griffiths and D. F. Schroeter, ``Introduction to Quantum Mechanics,'' Cambridge University Press, 3rd edition, 2017.

\bibitem{Bohm2003} A. Bohm, A. Mostafazadeh, H. Koizumi, Q. Niu, and J. Zwanziger, ``The Geometric Phase in Quantum Systems: Foundations, Mathematical Concepts, and Applications in Molecular and Condensed Matter Physics,'' Springer, 2003.

\bibitem{Shen2017} S.-Q. Shen, ``Topological Insulators: Dirac Equation in Condensed Matter,'' Springer, 2nd edition, 2017.

\end{thebibliography}

\section{Future Work}

Future work could explore:

\begin{itemize}
    \item More extensive mapping of the parameter space to identify all possible topological phases
    \item Deeper analysis of the physical meaning of the observed topological properties
    \item Application to real quantum systems and materials
    \item Extension to higher-dimensional parameter spaces
\end{itemize}

\end{document}
